M.Sc. Proposal: \\ \bigskip Computational Electrokinetics

### 2013-02-28 notes/thesis-latex-template/thesis.tex
Update thesis.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Electrophoresis for large $\beta$ }
When the electric field is large, nonlinear effects dominate the solution.
In order to investigate the solution for strong electric fields, 
we have run the solver on the TAMNUN HPC cluster \cite{tamnun}.
Large grid sizes are used to verify that the numerical solution converges when the grid becomes finer and that no numerical artifacts are present in the solution.
Moreover, several $R_{max} \in \{10, 30, 100\}$ values are used, to make sure it has no major impact on the solution.
The results are shown in the following figures. 
The steady-state velocity results are shown for several grid resolutions and several $R_{max}$ values in Figure \ref{fig:LargeBetaV_grids}, and the estimated rate of convergence is shown in Figure \ref{fig:LargeBetaRatio}.

@@

The streamline pattern does not change drastically for different electric field magnitudes. 
The resulting fluid flow is nearly proportional to that of the Stokes creeping flow, 
as shown in Figure \ref{fig:LargeBeta_Psi}. 
The difference between $\Psi(r, \theta)$ and $\Psi(r, \pi-\theta)$ is shown in Figure \ref{fig:LargeBeta_PsiAsymm}, demonstrating that the flow is almost symmetric (with relative
difference up to approximately $0.5\%$) for $\beta < 5$.

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
\section{Discussion} \label{sec:discussion}
A numerical framework for a
nonlinear electrokinetic transport system is described.
The framework is used to construct numerical solvers 
for two electrokinetic problems: the migration of ion-exchangers
and surface-charged inert particle electrophoresis.
The numerical results for weak electric field 
exhibit a good correspondence to the asymptotic models as shown above.
Two interesting phenomena (cubic regime and downstream vortex) were predicted for the ion-exchanger
using the numerical results, and were verified using asymptotic expansion of the nonlinear system.
The electrophoresis problem was solved on a high-resolution grid to validate the asymptotic 
cubic correction of the steady-state velocity, as well as providing insights of the solution behavior 
for moderate and strong electric fields.
For strong electric field, the nonlinear effects become dominant and
a boundary-layer structure begins to form.
Due to diminishing ion concentration $C$ at the front of the particle,
a boundary layer of thickness approximately $0.2 e^{-0.5\beta}$ 
is formed with respect to the electric field $\bE$ at the front of the particle.
The value of $E_r$ at ($r = 1$, $\theta = 0$) is given approximately by 
$e^{\beta/2}$, while the value of $C(r = 1, \theta = 0)$ is given approximately 
by $e^{-\beta/2}$, and their product remains close to 1 for all $\beta$ that we could compute.
However, the velocity of the particle remains nearly linear with $\beta$
and the velocity field is very close to being symmetric.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Electrophoresis for large $\beta$ }
When the electric field is large, nonlinear effects dominate the solution.
In order to investigate the solution for strong electric fields, 
we have run the solver on the TAMNUN HPC cluster \cite{tamnun}.
Large grid sizes are used to verify that the numerical solution converges when the grid becomes finer and that no numerical artifacts are present in the solution.
Moreover, several $R_{max} \in \{10, 30, 100\}$ values are used, to make sure it has no major impact on the solution.
The results are shown in the following figures. 
The steady-state velocity results are shown for several grid resolutions and several $R_{max}$ values in Figure \ref{fig:LargeBetaV_grids}, and the differences between the numerical results for
steady-state velocity and the linear regime are shown in Figure \ref{fig:LargeBetaV_DeltaFromLinear_grids}.
The estimated rate of convergence is shown in Figure \ref{fig:LargeBetaRatio}.

@@

The streamline pattern does not change drastically for different electric field magnitudes. 
The resulting fluid flow is nearly proportional to that of the Stokes creeping flow, 
as shown in Figure \ref{fig:LargeBeta_Psi} and \ref{fig:LargeBeta_PsiMinusUniformFlow}. 
The difference between $\Psi(r, \theta)$ and $\Psi(r, \pi-\theta)$ is shown in Figure \ref{fig:LargeBeta_PsiAsymm}, demonstrating that the flow is almost symmetric (with relative
difference up to approximately $0.5\%$) for $\beta < 5$.

@@

\begin{figure}
    \begin{center}
    \includegraphics[width=.45\textwidth]{figs/PsiMinusUniformFlow/sol_beta=1.000e-01_[513x513]_Rmax=10.0_Du=1.00_zeta=10.00_alpha=0.50_[1]_Psi.eps}
    \includegraphics[width=.45\textwidth]{figs/PsiMinusUniformFlow/sol_beta=8.000e-01_[513x513]_Rmax=10.0_Du=1.00_zeta=10.00_alpha=0.50_[1]_Psi.eps}
    \includegraphics[width=.45\textwidth]{figs/PsiMinusUniformFlow/sol_beta=5.000e+00_[513x513]_Rmax=10.0_Du=1.00_zeta=10.00_alpha=0.50_[1]_Psi.eps}
    \includegraphics[width=.45\textwidth]{figs/PsiMinusUniformFlow/LinearStokes_Psi-uniformflow.eps}
        \caption[Streamfunction results]
        {$\beta = 0.1, 0.8, 5.0$ and Stokes streamlines results after subtraction of the uniform
        flow streamfunction ($\frac{\cU}{2}r^2 \sin^2\theta$): grid size $512 \times 512$, 
        $R_{max} = 30$, $\Du = 1$, $\bar\zeta = 10$, $\alpha = 0.5$.
        The numerical results for $\Psi$ is scaled by $\beta^{-1}$, so the linear regime
        will appear the same for different $\beta$ values.
        Even for the largest beta we could compute, the velocity remains fairly symmetric and 		
        "uneventful" also in the framework of the "laboratory" rather than that of the particle.}
	    \label{fig:LargeBeta_PsiMinusUniformFlow}	    
    \end{center}
\end{figure}

@@

\begin{figure}
    \begin{center}
    \includegraphics[width=.6\textwidth]{figs/C_Er_product.eps}
        \caption{The product of $C(r=1, \theta=0) E_r(r=1, \theta=0) / \beta$ as
        a function of $\beta$.}
	    \label{fig:C_Er_product}	    
    \end{center}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
\section{Discussion} \label{sec:discussion}
A numerical framework for a
nonlinear electrokinetic transport system is described.
The framework is used to construct numerical solvers 
for two electrokinetic problems: the migration of ion-exchangers
and surface-charged inert particle electrophoresis.
The numerical results for weak electric field 
exhibit a good correspondence to the asymptotic models as shown above.
Two interesting phenomena (cubic regime and downstream vortex) were predicted for the ion-exchanger
using the numerical results, and were verified using asymptotic expansion of the nonlinear system.
The electrophoresis problem was solved on a high-resolution grid to validate the asymptotic 
cubic correction of the steady-state velocity, as well as providing insights of the solution behavior 
for moderate and strong electric fields.
For strong electric field, the nonlinear effects become dominant and
a boundary-layer structure begins to form.
Due to diminishing ion concentration $C$ at the front of the particle,
a boundary layer of thickness approximately $0.2 e^{-0.5\beta}$ 
is formed with respect to the electric field $\bE$ at the front of the particle.
The value of $E_r / \beta$ at ($r = 1$, $\theta = 0$) is given approximately by 
$e^{\beta/2}$, while the value of $C(r = 1, \theta = 0)$ is given approximately 
by $e^{-\beta/2}$, and their product remains close to 1 for all $\beta$ that we could compute
(as shown in Figure \ref{fig:C_Er_product}).
However, the velocity of the particle remains nearly linear with $\beta$
and the velocity field is very close to being symmetric.
